---
title: "project5_in_R"
author: "Isaac Moore"
date: "9/4/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setting up environment}
setwd("~/Google Drive/data_science/general_assembly/Projects/DSI_SM_Project5/r")
library(tidyverse)
library(RPostgreSQL)
library(stringr)
```

```{r adding table to titanic database}
# had to add a new table of the test dataframe, since someone had deleted it table form the database.

# test_clean <- read_csv("../data/test.csv")
# dbWriteTable(con, "titanic_test",
#             value = test_clean, append = FALSE, row.names = FALSE)
```

# Part 1: Aquire the Data
### 1. Connect to the remote database
```{r reading in data, include=TRUE}
pw <- {
  "gastudents"
}
 
# loads the PostgreSQL driver
drv <- dbDriver("PostgreSQL")
# creates a connection to the postgres database
# note that "con" will be used later in each connection to the database
con <- dbConnect(drv, dbname = "titanic",
                 host = "dsi.c20gkj5cvu3l.us-east-1.rds.amazonaws.com", port = 5432,
                 user = "dsi_student", password = pw)
rm(pw) # removes the password
```

### 2. Query the database and aggregate the data
```{r get data from database, include=TRUE}
train <- dbGetQuery(con, "SELECT * from titanic_train")
test <- dbGetQuery(con, "SELECT * from titanic_test")
```

# Part 2: Exploratory Data Analysis
### 1. Describe the Data
```{r create a copy of the "train" dataframe to "df"}
df <- train
```

```{r checking for null values}
colSums(is.na(df))
```

```{r Sumary Statistics for each column}
summary(df)
```
Exploring survival statistics
```{r explore data types in each column}
glimpse(df)
```

```{r how many survived?}
df$Survived <- as.factor(df$Survived)
levels(df$Survived) <- c("Victims", "Survived")
table(df$Survived)
```

```{r how many survived per each Sex}
df %>% group_by(Sex, Survived) %>% summarise(n = n())
```

```{r visualize how many survived per each Sex}
df %>% group_by(Sex, Survived) %>% 
        summarise(n = n()) %>%
        ggplot(aes(x = Sex, y = n)) + 
                geom_bar(stat = "identity") + 
                facet_grid(.~Survived) + 
                labs(title = "Survival counts per gender", x = "Sex", y = "Count")
```
More females survived, than perished, in our training dataset. 


```{r histogram of age per survival}
df %>% ggplot(aes(x = Age)) + 
                geom_histogram(binwidth = 5) + 
                facet_grid(.~Survived) + 
                geom_vline(xintercept = mean(df$Age, na.rm = T),colour = "red", show.legend = TRUE) + 
                labs(title = "Histogram of Age per survival", y = "Count", x = "Age")
```
If you are younger you were more likely to survive.

```{r Average ticket price paid per gender in each class}
df %>% group_by(Sex, Pclass) %>% 
        summarise(price = mean(Fare)) %>%
        ggplot(aes(y = price, x = Pclass, col = factor(Sex))) +
                geom_bar(stat = "identity", position = "dodge") + 
                labs(title = "Average prices paid per class for Male/Female", x = "Passenger Class", y = "Average Ticket Price", color = "Sex")
```
Females on average paid more than males, especially in first class.

```{r extract title and add to a new column}
df$title <- str_extract(df$Name, regex("[A-Z]\\w+\\."))
df$title[is.na(df$title)] <- "Other"
table(df$title)
```

```{r average age of survival per title}
ggplot(df, aes(x = title, y = mean(Age, na.rm = T), col = Survived)) + 
        geom_bar(stat = "identity", position = "stack") + 
        labs(title = "Average age of survival per title", y = "Average Age", x = "Title") + 
        coord_flip()
```
Unmarried women (Miss.) had a better survival rate (per average age) vs married women (Mrs.)

# Part 3: Data Wrangling
### 1. Create Dummy Variables for Sex
**I will convert the Sex column to a factor, which will work better in R ** 
```{r convert sex and passesnger class columns to a factor}
df$Sex <- as.factor(df$Sex)
df$Pclass <- as.factor(df$Pclass)
df$Embarked <- as.factor(df$Embarked)
```

Fill NA values...
```{r fill NA values}
df$Age[is.na(df$Age)] <- mean(df$Age, na.rm = T)# Filling with the mean Age
df$Cabin[is.na(df$Cabin)] <- "???" # too many to drop the columns, filling with '???'
df <- na.omit(df)
colSums(is.na(df))
```



# Part 4: Logistic Regression and Model Validation
### 1. Define the variables that we will use in our classification analysis  
**We will be using the *Pclass + Sex + Age + Parch + Fare + Embarked* columns from the dataframe to predict who survived on the Titanic**  

### 2. Transform "Y" into a 1-Dimensional Array for SciKit-Learn  
**No need to perform for logistic regression in R, you are able to specify our dependent and independent variables in the call to formulate the model.**  

### 3. Conduct the logistic regression  
```{r creating the model}
model <- glm(Survived ~ Pclass + Sex + Age + Parch + Fare + Embarked, family=binomial(link='logit'), data=df)
```
### 4. Examine the coefficients to see our correlations
```{r examine the coefficients}
summary(model)
```
### 6. Test the Model by introducing a Test or Validaton set
```{r convert the test columns to factors}
test$Sex <- as.factor(test$Sex)
test$Pclass <- as.factor(test$Pclass)
test$Embarked <- as.factor(test$Embarked)
```


```{r predict survival on the test set}
test_sub <- test %>% select(Pclass, Sex, Age, Parch, Fare, Embarked)
preds <- predict(model,test_sub,type='response')
preds[1:10]
```



